\documentclass[../main]{subfiles}
\chapter{Obstructions}
\begin{document}
\section{The Oriented Universal Bundle}
Let ${\xtilde \grassmannian}_n({\mathbb R}^{n + k})$ denote the Grassmann manifold consisting of all \defemph{oriented} $n$ planes in $(n + k)$--space. Just as in \S\ref{ch:5}, this can be topologized as a quotient space of the Stiefel manifold $\StiefelManifold_n({\mathbb R}^{n + k})$. Clearly ${\xtilde \grassmannian}_n({\mathbb R}^{n + k})$ is a $2$--fold covering space of the unoriented Grassmann manifold $\grassmannian_n({\mathbb R}^{n + k})$. It is easy to check that ${\xtilde \grassmannian}_n({\mathbb R}^{n + k})$ is a compact CW--complex of dimension $n k$. Passing to the direct limit as $k \to \infty$, we obtain an infinite CW--complex ${\xtilde \grassmannian}_n = {\xtilde \grassmannian}_n({\mathbb R}^\infty)$. (The notations $\BSO(n)$, respectively $\BO(n)$, are often used for these spaces ${\xtilde \grassmannian}_n$ and $\grassmannian_n$.)\index{BO(n),BSO(n)@$\BO(n),\BSO(n)$}\index{Grassmannian manifold!\indexline oriented $\xtilde \grassmannian$}

The universal bundle $\gamma^n$ over $\grassmannian_n$ lifts to an oriented $n$--plane bundle over ${\xtilde \grassmannian}_n$. We will denote this \defemph{oriented universal bundle}\index{cannonical bundle!\indexline oriented $\xtilde{\tautological}^n$} by the symbol ${\xtilde \gamma}^n$. It is clear that for any oriented $n$--plane bundle $\xi$, each bundle map $\xi \longrightarrow \gamma^n$ lifts uniquely to an orientation preserving bundle map $\xi \longrightarrow {\xtilde \gamma}^n$. \index{universal bundle!\indexline oriented}

The mod $2$ cohomology of ${\xtilde \grassmannian}_n$ can be computed as follows. (Compare \S\ref{ch:7}.)

\begin{theorem}
\label{thm:12.4}\index{cohomology!\indexline of $\xtilde{\grassmannian}_n$}
The cohomology $\homology^\ast({\xtilde \grassmannian}_n; {\mathbb Z}/2)$ is a polynomial algebra over ${\mathbb Z}/2$, freely generated by the Stiefel--Whitney classes $\sw_2({\xtilde \gamma}^n)$, $\ldots$, $\sw_n({\xtilde \gamma}^n)$.
\end{theorem}

In particular the group $\homology^1({\xtilde \grassmannian}_n; {\mathbb Z}/2)$ is zero. It follows that $\sw_1({\xtilde \gamma}^n) = 0$, and hence that $\sw_1(\xi) = 0$ for any oriented vector bundle $\xi$ over a para--compact base space. (Compare Problem \ref{prob:12.A}.)

\begin{proof}[Proof of \ref{thm:12.4}]
By \ref{cor:12.3} there is an exact sequence \[\ldots \varrightarrow{} \homology^{j - 1}(\grassmannian_n) \varrightarrow{\smile c} \homology^j(\grassmannian_n) \varrightarrow{p^\ast} \homology^j({\xtilde \grassmannian}_n) \longrightarrow \ldots,\] where $c$ is the first Stiefel--Whitney class of the line bundle associated with the $2$--fold covering, and where $p : {\xtilde \grassmannian}_n \longrightarrow \grassmannian_n$ is the natural map. This class $c$ cannot be zero. For otherwise the sequence \[0 \varrightarrow{} \homology^0(\grassmannian_n) \varrightarrow{} \homology^0({\xtilde \grassmannian}_n) \varrightarrow{} \homology^0(\grassmannian_n) \varrightarrow{\smile c} \ldots\] would imply that ${\xtilde \grassmannian}_n$ had two components, contradicting the evident fact that any oriented $n$--plane in ${\mathbb R}^\infty$ can be deformed continuously to any other oriented $n$--plane. Thus $c = \sw_1(\gamma^n)$, using \S\ref{thm:07.01}, and a straightforward argument completes the proof. 
\end{proof}
\end{document}