\chapter{A Cell Structure for Grassmann Manifolds}



This section will describe a canonical cell subdivision, due to Ehresmann \cite{85}, which makes the infinite Grassmann manifold $\grass{n}{\infty}$ into a CW-complex. Each finite Grassmann manifold $\grass{n}{n+k}$ appears as a finite subcomplex. This cell structure has been used by Pontrjagin \cite{69} and by Chern \cite{70} as a basis for the theory of characteristic classes. The reader should consult these sources, as well as \cite{55} for further information. For a thorough treatment of cell complexes in general, consult \cite{71}. Grassmann manifolds appear there on p. 17.

First recall some definitions. Let $\disk^{p}$ denote the \textit{unit disk} in $\R^p$, consisting of all vectors $v$ with $| v | \leq 1$. The \textit{interior} of $\disk^{p}$ is defined to be the subset consisting of all $v$ with $|v|<1$. For the special case $p=0$, both $\disk^{p}$ and its interior consist of a single point.

Any space homeomorphic to $\disk^{p}$ is called a \textit{closed $p$-cell}; and any space homeomorphic to the interior of $\disk^{p}$ is called an \textit{open $p$-cell}. For example $\R^p$ is an open $p$-cell.

\begin{definition}[J. H. C. Whitehead, 1949]\label{def:6-1}
	A \textit{CW-complex} consists of a Hausdorff space $K$, called the \textit{underlying space}, together with a partition of $K$ into a collection $\{e_{\alpha}\}$ of disjoint subsets, such that four conditions are satisfied:
	\begin{enumerate}[label=\arabic*)]
		\item Each $e_{\alpha}$ is topologically an open cell of dimension $n(\alpha) \geq 0$. Furthermore for each cell $e_{\alpha}$ there exists a continuous map
\[
\map{f}{\disk^{n(\alpha)}}{K}f
\]
which carries the interior of the disk $\disk^{n(\alpha)}$ homeomorphically onto $e_{a}$. (This $f$ is called a \textit{characteristic map} for the cell $e_{\alpha}$.)
\item Each point $x$ which belongs to the closure $\overline{e}_{\alpha}$, but not to $e_{\alpha}$ itself, must lie in a cell $e_\beta$ of lower dimension.
	\end{enumerate}

If the complex is \textit{finite} (i.e., if there are only finitely many $e_{a}$ ), then these two conditions suffice. However in general two further conditions are needed. A subset of $K$ is called a [finite] \textit{subcomplex} if it is a closed set and is a union of [finitely many] $e_{a}$ 's.

\begin{enumerate}[label=\alph*),leftmargin=2\parindent ]
	\setcounter{enumi}{2}
	\item \textsc{Closure finiteness.} Each point of $K$ is contained in a finite subcomplex.
	
	\item \textsc{Whitehead topology.} $K$ is topologized as the direct limit of its finite subcomplexes. I.e., a subset of $K$ is closed if and only if its intersection with each finite subcomplex is closed.
	\end{enumerate}
Note that the closure $\bar{e}_{\alpha}$ of a cell of $K$ need not be a cell. For example the sphere $\Sphere{n}$ can be considered as a CW-complex with one $0$-cell and one $n$-cell. In this case the closure of the $n$-cell is equal to the entire sphere.
\end{definition}
A theorem of Miyazaki \cite{72} asserts that every CW-complex is paracompact. (Compare \cite[p.~419]{73}.)

The cell structure for the Grassmann manifold $\grass{n}{m}$ is obtained as follows. Recall that $\R^m$ contains subspaces
\[
\R^{0} \subset \R^{1} \subset \R^{2} \subset\dots \subset \R^m;
\]
where $\R^{k}$ consists of all vectors of the form $v=(v_{1},\dots, v_{k}, 0,\dots, 0)$. Any $n$-plane $X \subset \R^m$ gives rise to a sequence of integers
\[
0 \leq \dim(X \cap \R^{1}) \leq \dim(X \cap \R^{2}) \leq\dots \leq \dim(X \cap \R^m)=n.
\]
Two consecutive integers in this sequence differ by at most $1$. This fact is proved by inspecting the exact sequence
\[
0 \xrightarrow{\qquad}X \cap \R^{k-1} \xrightarrow{\qquad}X \cap \R^{k}\xrightarrow{\ k\text{-th coordinate }}  \R.
\]
Thus the above sequence of integers contains precisely $n$ ``jumps" By a Schubert symbol $\sigma=(\sigma_{1},\dots, \sigma_{n})$ is meant a sequence of $n$ integers satisfying
\[
1 \leq \sigma_{1}<\sigma_{2}<\dots<\sigma_{n} \leq m.
\]
For each Schubert symbol $\sigma$, let $e(\sigma) \subset \grass{n}{m}$ denote the set of all $n$-planes $X$ such that
\[
\dim(X \cap \R^{\sigma_{i}})=i,\quad \dim(X \cap \R^{\sigma_{i}-1})=i-1
\]
for $i=1,\dots, n$. Evidently each $X \in \grass{n}{m}$ belongs to precisely one of the sets  $e(\sigma)$. We will see presently that $e(\sigma)$ is an open cell\footnote{The closure $\overline{e}(\sigma)$ is called a Schubert variety. (Compare [Schubert].) In the notation of Chern and Wu, the cell $e(\sigma)$ is indexed not by the sequence $\sigma=(\sigma_{1},\dots, \sigma_{n})$ but rather by the modified sequence $(\sigma_{1}-1, \sigma_{2}-2,\dots, \sigma_{n}-n)$, which is more convenient to use for many purposes.} of dimension $d(\sigma)=(\sigma_{1}-1)+(\sigma_{2}-2)+\dots+(\sigma_{n}-n)$.

Let $\mathbb{H}^{k} \subset \R^{k}$ denote the open half-space consisting of all $x= (\xi_{1},\dots, \xi_{k}, 0,\dots, 0)$ with $\xi_{k}>0 $. Note that an $n$-plane $X$ belongs to $e(\sigma)$ if and only if it possesses a basis $x_{1},\dots, x_{n}$ so that
\[
x_{1} \in \mathbb{H}^{\sigma_{1}},\dots, x_{n} \in \mathbb{H}^{\sigma_{n}}.
\]
For if $X$ possesses such a basis, then the exact sequence above shows that
\[
\dim(X \cap \R^{\sigma_{i}})>\dim(X \cap \R^{\sigma_{i}-1})
\]
for $i=1,\dots, n$, hence $X \in e(\sigma) $. The converse is proved similarly. In terms of matrices, the $n$-plane $X$ belongs to $e(\sigma)$ if and only if it can be described as the row space of an $n \times m$ matrix $\left[x_{ij}\right]$ of the form

\[\begin{bmatrix}
	* & \dots & *10 & \dots & 000 & \dots & 000 & \dots & 0 \\
	* & \dots & *** & \dots & *10 & \dots & 000 & \dots & 0 \\
	\vdots &  & \vdots &  & \vdots &  & \vdots &  & \vdots \\
	* & \dots & *** & \dots & *** & \dots & *10 & \dots & 0
\end{bmatrix}\]
where the $i$-th row has $\sigma_{i}$-th entry positive (say equal to $1$), and all subsequent entries zero.
\begin{lemma}\label{lem-6-2}
	Each $ n $-plane $X \in e(\sigma)$ possesses a unique orthonormal basis $(x_{1},\dots, x_{n})$ which belongs to $\mathbb{H}^{\sigma_{1}} \times\dots \times \mathbb{H}^{\sigma_{n}}$.
\end{lemma}
\begin{proof}
	The vector $x_{1}$ is required to lie in the $1$-dimensional vector space $X \cap \R^{\sigma_{1}}$, and to be a unit vector. This leaves only two possibilities for $x_{1}$, and the condition that the $\sigma_{1}$-th coordinate be positive specifies one of these two. Now $x_{2}$ is required to be a unit vector in the $2$ dimensional space $X\cap\R^{\sigma_{2}}$, and to be orthogonal to $x_{1} $. Again this leaves two possibilities, and the condition that the $\sigma_{2}$-th coordinate be positive specifies one of these two. Continuing by induction, it follows that $x_{3}, x_{4},\dots, x_{n}$ are also uniquely determined.
\end{proof}
\begin{definition}\label{def:6-2}
	Let $e^{\prime}(\sigma)=V_{n}^{0}(\R^m) \cap(\mathbb{H}^{\sigma_{1}} \times\dots \times \mathbb{H}^{\sigma_{n}})$ denote the set of all orthonormal $n$-frames $(x_{1},\dots, x_{n})$ such that each $x_{i}$ belongs to the open half-space $\mathbb{H}^{\sigma_i}$. Let $\overline{e}^{\prime}(\sigma)$ denote the set of orthonormal frames $(x_{1},\dots, x_{n})$ such that each $x_{i}$ belongs to the closure $\overline{\mathbb{H}}^{\sigma_{i}}$.
\end{definition}
\begin{lemma}\label{lem-6-3}
	The set $\overline{e}^{\prime}(\sigma)$ is topologically a closed cell of dimension $d(\sigma)=(\sigma_{1}-1)+(\sigma_{2}-2)+\dots+(\sigma_{n}^{-n})$, with interior $e^{\prime}(\sigma) $. Furthermore $q$ maps the interior $e^{\prime}(\sigma)$ homeomorphically onto $e(\sigma)$.
\end{lemma}

Thus $e(\sigma)$ is actually an open cell of dimension $d(\sigma)$. Furthermore the map
\[
\map{q \mid_ {\overline{e}^{\prime}(\sigma)}}{\overline{e}^{\prime}(\sigma)}{\grass{n}{m}}
\]
will serve as a characteristic map for this cell.
\begin{proof}
The proof of \cref{lem-6-3} will be by induction on $n$. For $n=1$ the set $\overline{e}^{\prime}(\sigma_{1})$ consists of all vectors
\[
x_{1}=(x_{11}, x_{12},\dots, x_{1 \sigma_{1}}, 0,\dots, 0)
\]
with $\Sigma x_{1 i}^{2}=1, x_{1 \sigma_{1}} \geq 0 $. Evidently $\overline{e}^{\prime}(\sigma_{1})$ is a closed hemisphere of dimension $\sigma_{1}-1$, and therefore is homeomorphic to the disk $\disk^{\sigma_{1}-1}$.

Given unit vectors $u, v \in \R^{m}$ with $u \neq-v$, let $T(u, v)$ denote the unique rotation of $\R^m$ which carries $u$ to $v$, and leaves everything orthogonal to $u$ and $v$ fixed. Thus $T(u, u)$ is the identity map and $T(v, u)=T(u, v)^{-1} .$ Alternatively $T(u, v)$ can be defined by the formula
\[
T(u, v) x=x-\frac{(u+v) \cdot x}{1+u \cdot v}(u+v)+2(u \cdot x) v
\]
In fact the function $T(u, v)$ defined in this way is linear in $x$, and has the correct effect on the vectors $u, v$, and on all vectors orthogonal to $u$ and $v$. It follows from this formula that:

\begin{enumerate}[label=\arabic*),leftmargin=2\parindent ]
	\item $T(u, v) x$ is continuous as a function of three variables; and
	
	\item if $u, v \in \R^{k}$ then $T(u, v) x \equiv x$ (modulo $\R^{k}$ ).
\end{enumerate}

Let $b_{i} \in \mathbb{H}^{\sigma_{i}}$ denote the vector with $\sigma_{i}$-th coordinate equal to $1$, and all other coordinates zero. Thus $(b_{1},\dots, b_{n}) \in e^{\prime}(\sigma) $. For any $n$-frame $(x_{1},\dots, x_{n}) \in \overline{e}^{\prime}(\sigma)$ consider the rotation
\[
T=T(b_{n}, x_{n}) \circ T(b_{n-1}, x_{n-1}) \circ\dots \circ T(b_{1}, x_{1})
\]
of $\R^m$. This rotation carries the $n$ vectors $b_{1},\dots, b_{n}$ to the vectors $x_{1},\dots, x_{n}$ respectively. In fact the rotations $T(b_{1}, x_{1}),\dots, T(b_{i-1}, x_{i-1})$ leave $b_{i}$ fixed (since $b_{i} \cdot b_{j}=b_{i} \cdot x_{j}=0$ for $i>\mathrm{j}$ ); the rotation $T(b_{i}, x_{i})$ carries $b_{i}$ to $x_{i} ;$ and the rotations $T(b_{i+1}, x_{i+1}),\dots, T(b_{n}, x_{n})$ leave $x_{i}$ fixed.

Given an integer $\sigma_{n+1}>\sigma_{n}$ let $d$ denote the set of all unit vectors $u \in \overline{\mathbb{H}}^{\sigma_{n+1}}$ with
\[
b_{1} \cdot u=\dots=b_{n} \cdot u=0.
\]
Evidently $d$ is a closed hemisphere of dimension $\sigma_{n+1}-n-1$, and hence is topologically a closed cell. We will construct a homeomorphism
\[\map{f}{\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n}) \times d}{\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n+1})}.
\]
In fact $f$ is defined by the formula
\[
f((x_{1},\dots, x_{n}), u)=(x_{1},\dots, x_{n}, Tu)
\]
where the rotation $T$ depends on $x_{1},\dots, x_{n}$, as above. To prove that $(x_{1},\dots, x_{n}, Tu)$ actually belongs to $\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n+1})$ we note that
\[
x_{i} \cdot Tu=\mathrm{Tb}_{i} \cdot Tu=b_{i} \cdot u=0
\]
for $i \leq n$, and that
\[
Tu \cdot Tu=u \cdot u=1
\]
where $Tu \in \overline{\mathbb{H}}^{\sigma_{n+1}}$ since $Tu \equiv u(\bmod \R^{\sigma_{n}})$. Evidently $f$ maps $\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n}) \times d$ continuously to $\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n+1}) $. Similarly the formula
\[
u=T^{-1} x_{n+1}=T(x_{1}, b_{1}) \circ\dots \circ T(x_{n}, b_{n}) x_{n+1} \in \mathrm{D}
\]
shows that $f^{-1}$ is well defined and continuous.

Thus $\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n+1})$ is homeomorphic to the product $\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n}) \times \mathrm{D}$ It follows by induction on $n$ that each $\overline{e}^{\prime}(\sigma)$ is a closed cell of dimension $d(\sigma) $. A similar induction shows that each $e^{\prime}(\sigma)$ is the interior of the cell $\overline{e}^{\prime}(\sigma) $. In fact the homeomorphism
\[
\map{f} {\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n}) \times d}{\overline{e}^{\prime}(\sigma_{1},\dots, \sigma_{n+1})}
\]
carries the product $e^{\prime}(\sigma_{1},\dots, \sigma_{n}) \times \mathrm{Int\ D}$ onto $e^{\prime}(\sigma_{1},\dots, \sigma_{n+1})$.

\noindent\textsc{Proof that the map}
$
\map{q\mid_ {e^{\prime}(\sigma)}}{ e^{\prime}(\sigma)}{e(\sigma)}
$
\textsc{is a homeomorphism.} According to \cref{lem-6-2}, $q$ carries $e^{\prime}(\sigma)$ in one-one fashion onto $e(\sigma) $. On the other hand, if $(x_{1},\dots, x_{n})$ belongs to the ``boundary" $\overline{e}^{\prime}(\sigma)-e^{\prime}(\sigma)$, then the $n$-plane $X=q(x_{1},\dots, x_{n})$ does not belong to $e(\sigma)$, for one of the vectors $x_{i}$ must lie in the boundary $\R^{\sigma_{i}-1}$ of the half-space $\overline{\mathbb{H}}^{\sigma_{i}}$. This implies that
\[
\dim(X \cap \R^{\sigma_{i}-1}) \geq i,
\]
and hence that $X \notin e(\sigma)$.

Now let $A \subset e^{\prime}(\sigma)$ be a relatively closed subset. Then $\overline{A} \cap e^{\prime}(\sigma)=A$, where the closure $\overline{A} \subset \overline{e}^{\prime}(\sigma)$ is compact, hence $q(\overline{A})$ is closed. The preceding paragraph implies that $q(\overline{A}) \cap e(\sigma)=q(A)$, and it follows that $q(A) \subset e(\sigma)$ is a relatively closed set. Thus $q$ maps the cell $e^{\prime}(\sigma)$ homeomorphically onto $e(\sigma)$.
\end{proof}
\begin{theorem}\label{thm-6-4}
	The $\binom{m}{n}$ sets $e(\sigma)$ form the cells of a CW-complex with underlying space $\grass{n}{m}$. Similarly taking the direct limit as $m \rightarrow \infty$, one obtains an infinite CW-complex with underlying space $\grass{n}{}=\grass{n}{\infty}$.
\end{theorem}
\begin{proof}
	We must first show that each point in the boundary of a cell $e(\sigma)$ belongs to a cell $e(\tau)$ of lower dimension. Since $\overline{e}^{\prime}(\sigma)$ is compact, the image $q \overline{e}^{\prime}(\sigma)$ is equal to $\overline{e}(\sigma)$. Hence every $n$-plane $X$ in the boundary $\overline{e}(\sigma)-e(\sigma)$ has a basis $(x_{1},\dots, x_{n})$ belonging to $\overline{e}^{\prime}(\sigma)-e^{\prime}(\sigma)$ Evidently the vectors $x_{1},\dots, x_{n}$ are orthonormal, with $x_{i} \in \R^{\sigma_{i}}$. It follows that $\dim(X \cap \R^{\sigma_{i}}) \geq i$ for each $i$, thus the Schubert symbol $(\tau_{1},\dots, \tau_{n})$ associated with $X$ must satisfy
	\[
	\tau_{1} \leq \sigma_{1},\dots, \tau_{n} \leq \sigma_{n}.
	\]
	As above, one of the vectors $x_{i}$ must actually belong to $\R^{\sigma_{i}-1} $; hence the corresponding integer $\tau_{i}$ must be strictly less than $\sigma_{i}$. Therefore $d(\tau)<d(\sigma)$. Together with \cref{lem-6-3}, this completes the proof that $\grass{n}{m}$ is a finite CW-complex.
	
	Similarly $\grass{n}{\infty}$ is a CW-complex. The closure finiteness condition is satisfied since each $X \in \grass{n}{\infty}$ belongs to a finite subcomplex $\grass{n}{m}$. The space $\grass{n}{\infty}$ has the direct limit topology by definition.
\end{proof} 

It is instructive to look at the special case $n=1$. \begin{corollary}\label{cor-6-5}
The infinite projective space $\rp{\infty}=\grass{1}{\infty}$ is a CW-complex having one $r$-cell $e(r+1)$ for each integer $r \geq 0$. The closure $\overline{e}(r+1) \subset \rp{\infty}$ is equal to the finite projective space $\rp{r}$.	
\end{corollary}
The proof is straightforward.

Now let us count the number of $r$-cells in $\grass{n}{m}$ for arbitrary $n $. It is convenient to introduce the language of partitions.

\begin{definition}\label{def:6-7}
A \defemph{partition of an integer} $r \geq 0$ is an unordered sequence $i_{1} i_{2} \dots i_{s}$ of positive integers with sum $\R$. The number of partitions of $r$ is customarily denoted by $p(r) $. Thus for $r \leq 10$ one has the following table.

\begin{table}[!h]
\centering\begin{tabular}{|l|lllllllllll|}
	\hline
	$r$ & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
	\hline
	$p(r)$ & 1 & 1 & 2 & 3 & 5 & 7 & 11 & 15 & 22 & 30 & 42 \\
	\hline
\end{tabular}
\end{table}	
\end{definition}

For example the integer 4 has five partitions, namely: 1111,112, 22,13, and $4 .$ The integer 0 has just one (vacuous) partition. (According to Hardy and Ramanujan the function $p(r)$ is asymptotic to $\exp (\pi \sqrt{2 r / 3}) / 4 r \sqrt{3}$ as $r\rightarrow \infty$. For further information see \cite{74}.)

To every Schubert symbol $(\sigma_{1},\dots, \sigma_{n})$ with $d(\sigma)=r$ and $\sigma_{n} \leq m$ there corresponds a partition $i_{1}\dots i_{s}$ of $r$, where $i_{1},\dots, i_{s}$ denotes the sequence obtained from $\sigma_{1}-1,\dots, \sigma_{n}-n$ by cancelling any zeros which may appear at the beginning of this sequence. Clearly
\[
1 \leq i_{1} \leq i_{2} \leq\dots \leq i_{s} \leq m-n
\]
and $s \leq n $. Thus

\begin{corollary}\label{cor-6-7}
The number of $r$-cells in $\grass{n}{m}$ is equal to the number of partitions of $r$ into at most $n$ integers each of which is $\leq m-n$.	
\end{corollary}
 In particular, if both $n$ and $m-n$ are $\geq r$, then the number of $r$-cells in $\grass{n}{m}$ is equal to $p(r)$.

Note that this corollary remains true if $m$ is allowed to take the value $+\infty$.

Here are five problems for the reader.

\begin{problem}\label{prob-6-A}
	Show that a CW-complex is finite if and only if its underlying space is compact.
\end{problem}

\begin{problem}\label{prob-6-B}
	Show that the restriction homomorphism
	\[\map{i^{*}}{\homology^{p}(\grass{n}{\infty})}{\homology^{p}(\grass{n}{n+k})}
	\]
	is an isomorphism for $p<k $. Any coefficient group may be used. (Compare the description of cohomology for CW-complexes in \Cref{app-A}.)
\end{problem}

\begin{problem}\label{prob-6-C}
	Show that the correspondence $x \stackrel{f}{\longrightarrow} \R^{1} \oplus X$ defines an embedding of the Grassmann manifold $\grass{n}{m}$ into $\grass{n+1}{}(\R^{1} \oplus \R^{m})=\grass{n+1}{m+1}$, and that $f$ is covered by a bundle map
	\[
	\mathcal{E}^{1} \oplus \gamma^{n}(\R^m) \longrightarrow \gamma^{n+1}(\R^{m+1}).
	\]
	Show that $f$ carries the $r$-cell of $\grass{n}{m}$ which corresponds to a given partition $i_{1}\dots i_{s}$ of $r$ onto the $r$-cell of $\grass{n+1}{m+1}$ which corresponds to the same partition $i_{1}\dots i_{s}$.
\end{problem}

\begin{problem}\label{prob-6-D}
	Show that the number of distinct Stiefel-Whitney numbers $w_{1}{ }^{r_{1}}\dots w_{n} { }^{r^{n}}[M]$ for an $n$-dimensional manifold is equal to $p(n)$.
\end{problem}
 
\begin{problem}\label{prob-6-E}
	Show that the number of $r$-cells in $\grass{n}{n+k}$ is equal to the number of $r$-cells in $\grass{k}{n+k}$ [or show that these two $CW$-complexes are actually isomorphic].
\end{problem}


%\section*{title}\label{sec5.8}