\documentclass[../main]{subfiles}

\begin{document}
\chapter*{Preface to the new typesetting}
\addcontentsline{toc}{chapter}{Preface To The New Typesetting}

The book was TeX'd up by the Texromancers, a latexing group. The credits for the typesetting of this book go to: Aareyan Manzoor, Abhishek Shivkumar, Yohan Wittgenstein, Kelvin Chan, RokettoJanpu, George Coote, John Cerkan, Mervyn Brumbach (III), and Carl Sun.  %put your names here.

Here is a link to a dyslexic friendly version: \url{https://aareyanmanzoor.github.io/assets/books/characteristic-classes-dyslexic.pdf} .

We added citations and references with hyperlinks. References to e.g. theorems/lemmas in the book are in blue, while citations to the bibliography is
in red. The bibliography also has URLs now, for easy access. Some of the books
in the bibliography had newer editions, so we went with those.

Some editor's notes were put in footnotes to point things out or talk about changes since the book came out.

A lot of little notation was changed to better fit the new latex'd version. For example, we changed tangent space to $\tangentspace{M}{x}$ rather than $D_x M$. Things like characteristic classes or homologies or special groups have formatting to emphasize them.

A section defining fibre bundles was added as the book uses the terminology in a few places. In the appendix, a section defining relative homology was added for the same reason.

Figures were redrawn to better fit the new style of the book.

\end{document}